---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)
```

# Word Vectors in R

<!-- badges: start -->
<!-- badges: end -->

The goal of this repository is to showcase how to pretrain word vectors and how to acces them in R.
This repository will not show you why you should use each of the methods or how they work.
Only how they are being used within the R language.
Examples will be focused on getting word vectors that will work well with [textrecipes](https://textrecipes.tidymodels.org/)'s [step_word_embeddings](https://textrecipes.tidymodels.org/reference/step_word_embeddings.html) step which requires a tibble where the first column contains tokens, and additional columns should contain embeddings vectors.

For more information about read:

- [Dive into Deep Learning: Natural Language Processing: Pretraining](http://d2l.ai/chapter_natural-language-processing-pretraining/index.html)

## Data and packages

We will be using the [hcandersenr](https://github.com/EmilHvitfeldt/hcandersenr) package to provide data for the examples.
In addition will be load [tidyverse](https://www.tidyverse.org/) and [tokenizers](https://github.com/ropensci/tokenizers).
These two packages are not essencial for this to work but provides a little ease of use.

```{r, message=FALSE}
library(hcandersenr)
library(tidyverse)
library(tokenizers)
```

We will start by collapsing the text from hcandersenr such that we have 1 fairy tale per string.

```{r}
fairy_tales <- hcandersen_en %>%
  nest(data = c(text)) %>%
  mutate(text = map_chr(data, ~ paste(.x$text, collapse = " "))) %>%
  pull(text)

str(fairy_tales)
```

Taking a look reveals we have a little under half a million tokens.
This is pretty good but properly on the low end for building good word vectors.

```{r}
fairy_tales %>% 
  count_words() %>% 
  sum()
```

## text2vec

### Package

https://github.com/bnosac/word2vec

### Fitting the model

```{r}
library(word2vec)

model <- word2vec(x = fairy_tales, type = "cbow", dim = 100, window = 5, iter = 20)
```

### Extraction word vectors

```{r}
word_vectors <- as.matrix(model) %>%
  as.data.frame() %>%
  rownames_to_column("tokens") %>%
  as_tibble()

word_vectors
```

## Glove

### Package

https://github.com/dselivanov/text2vec

### Fitting the model

```{r include=FALSE}
library(text2vec)

tokens <- space_tokenizer(fairy_tales)
it <- itoken(tokens, progressbar = FALSE)
vocab <- create_vocabulary(it)
vocab <- prune_vocabulary(vocab, term_count_min = 5L)
vectorizer <- vocab_vectorizer(vocab)
tcm <- create_tcm(it, vectorizer, skip_grams_window = 5L)
glove <- GlobalVectors$new(rank = 100, x_max = 10)
wv_main <- glove$fit_transform(tcm, n_iter = 10, convergence_tol = 0.01)
```

```{r eval=FALSE}
library(text2vec)

tokens <- space_tokenizer(fairy_tales)
it <- itoken(tokens, progressbar = FALSE)
vocab <- create_vocabulary(it)
vocab <- prune_vocabulary(vocab, term_count_min = 5L)
vectorizer <- vocab_vectorizer(vocab)
tcm <- create_tcm(it, vectorizer, skip_grams_window = 5L)
glove <- GlobalVectors$new(rank = 100, x_max = 10)
wv_main <- glove$fit_transform(tcm, n_iter = 10, convergence_tol = 0.01)
```
### Extraction word vectors

```{r inc}
wv_context <- glove$components
word_vectors <- wv_main + t(wv_context)

word_vectors <- word_vectors %>%
  as.data.frame() %>%
  rownames_to_column("tokens") %>%
  as_tibble()

word_vectors
```

## Fasttext

### Package

https://github.com/pommedeterresautee/fastrtext

### Fitting the model

```{r}
library(fastrtext)

texts <- tolower(fairy_tales)
tmp_file_txt <- tempfile()
tmp_file_model <- tempfile()
writeLines(text = texts, con = tmp_file_txt)
execute(commands = c("skipgram", "-input", tmp_file_txt, "-output", tmp_file_model, 
                     "-verbose", 1, "-dim", 100))

model <- load_model(tmp_file_model)
```

### Extraction word vectors

```{r}
word_vectors <- get_word_vectors(model) %>%
  as.data.frame() %>%
  rownames_to_column("tokens") %>%
  as_tibble()

word_vectors
```
